---
title: "TSAR: Clinical EDA 2"
author: "Joshua Burkhart"
date: "August 18, 2016"
output: 
  pdf_document: 
    fig_width: 9
    fig_height: 6
    latex_engine: xelatex
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      results='hide', # markup marks up the results; asis return texts as-is; hide hides the results
                      comment=NA,
                      message=FALSE,
                      warning=FALSE,
                      error=FALSE,
                      fig.path="Figs/"
                      )
```

```{r libs, echo=FALSE, include=FALSE}
library(magrittr)
library(ggplot2)
library(dplyr)
library(memisc)
library(pander)
```

```{r globals}
# Assume working directory matches source file location
# R Studio: Session->Set Working Directory->To Source File Location
#setwd("/media/burkhart/Media/Software/TSAR/src")
DATA_DIR <- "/media/burkhart/D052-7853/McWeeney/TSAR/"
TRAIN_SYMPTOM <- paste(DATA_DIR,"ViralChallenge_training_SymptomScoresByDay.tsv",sep="")
TRAIN_CLINICL <- paste(DATA_DIR,"ViralChallenge_training_CLINICAL.tsv",sep="")
PHAS1_CLINICL <- paste(DATA_DIR,"ViralChallenge_test_Phase1_CLINICAL.tsv",sep="")
PHAS2_CLINICL <- paste(DATA_DIR,"ViralChallenge_test_Phase2_CLINICAL.tsv",sep="")
PHAS3_CLINICL <- paste(DATA_DIR,"ViralChallenge_test_Phase3_CLINICAL.tsv",sep="")
```
  
```{r misc_options}
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))
```
  
```{r helper_functions}
# see http://stackoverflow.com/questions/33524669/labeling-outliers-of-boxplots-in-r
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x >= quantile(x, 0.75) + 1.5 * IQR(x))
}

add_virus_col <- function(x){
  x %>%
  dplyr::mutate(VIRUS = 
                  ifelse(grepl(" H1N1$",STUDYID) %>% as.character(),"H1N1",
                  ifelse(grepl(" H3N2$",STUDYID) %>% as.character(),"H3N2",
                  ifelse(grepl(" RSV$",STUDYID) %>% as.character(),"RSV",
                  ifelse(grepl("^Rhinovirus ",STUDYID) %>% as.character(),"Rhinovirus",
                  NA))))) %>%
    dplyr::mutate(VIRUS = VIRUS %>% as.factor()) %>%
  return()
}

age_by_study <- function(x,label){
  x %>%
  dplyr::group_by(STUDYID) %>%
  dplyr::mutate(outlier = ifelse(is_outlier(AGE), AGE, as.numeric(NA))) %>%
  ggplot(aes(x = STUDYID, y = AGE)) +
  geom_boxplot() +
  geom_text(aes(label = AGE), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Age",
       title=paste("Patient Age Distribution by Study ID:\n",
       label," Data",sep=""))
}

age_by_virus <- function(x,label){
  x %>%
  dplyr::group_by(VIRUS) %>%
  dplyr::mutate(outlier = ifelse(is_outlier(AGE), AGE, as.numeric(NA))) %>%
  ggplot(aes(x = VIRUS, y = AGE)) +
  geom_boxplot() +
  geom_text(aes(label = AGE), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Virus",
       y="Patient Age",
       title=paste("Patient Age Distribution by Virus:\n",
       label," Data",sep=""))
}

log_symptom_score_by_study <- function(x,label){
  x %>%
  dplyr::group_by(STUDYID) %>%
  dplyr::mutate(outlier = ifelse(is_outlier(LOGSYMPTSCORE_SC3), LOGSYMPTSCORE_SC3, as.numeric(NA))) %>%
  ggplot(aes(x = STUDYID, y = LOGSYMPTSCORE_SC3)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Log Symptom Score",
       title=paste("Patient Log Symptom Score Distribution by Study ID:\n",
       label," Data",sep=""))
}

log_symptom_score_by_virus <- function(x,label){
  x %>%
  dplyr::group_by(VIRUS) %>%
  dplyr::mutate(outlier = ifelse(is_outlier(LOGSYMPTSCORE_SC3), LOGSYMPTSCORE_SC3, as.numeric(NA))) %>%
  ggplot(aes(x = VIRUS, y = LOGSYMPTSCORE_SC3)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Log Symptom Score",
       title=paste("Patient Log Symptom Score Distribution by Virus:\n",
       label," Data",sep=""))
}

log_symptom_score_by_timehours_by_study <- function(x,label){
  x %>%
    dplyr::mutate(TIMEHOURS = as.factor(TIMEHOURS)) %>%
    dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
    dplyr::mutate(outlier = ifelse(is_outlier(JACKSON_SCORE), JACKSON_SCORE, as.numeric(NA))) %>%
    dplyr::arrange(TIMEHOURS) %>%
  ggplot(aes(x = TIMEHOURS, y = JACKSON_SCORE)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Jackson Score",
       title=paste("Patient Jackson Score Distribution by Study ID:\n",
       label," Data",sep="")) +
    facet_wrap(~STUDYID)
}

log_symptom_score_by_timehours_by_virus <- function(x,label){
  x %>%
    dplyr::mutate(TIMEHOURS = as.factor(TIMEHOURS)) %>%
        dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
    dplyr::mutate(outlier = ifelse(is_outlier(JACKSON_SCORE), JACKSON_SCORE, as.numeric(NA))) %>%
    dplyr::arrange(TIMEHOURS) %>% 
  ggplot(aes(x = TIMEHOURS, y = JACKSON_SCORE)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Jackson Score",
       title=paste("Patient Jackson Score Distribution by Virus:\n",
       label," Data",sep="")) +
    facet_wrap(~VIRUS)
}

symptom_score_by_studyday_by_study <- function(x,label){
  x %>%
    dplyr::mutate(STUDYDAY = as.factor(STUDYDAY)) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE)) %>%
    dplyr::mutate(outlier = ifelse(is_outlier(SUM_SCORE), SUM_SCORE, as.numeric(NA))) %>%
    dplyr::arrange(STUDYDAY) %>%
  ggplot(aes(x = STUDYDAY, y = SUM_SCORE)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Symptom Score Sum",
       title=paste("Patient Symptom Score Sum Distribution by Study ID:\n",
       label," Data",sep="")) +
    facet_wrap(~STUDYID)
}

symptom_score_by_studyday_by_virus <- function(x,label){
  x %>%
    dplyr::mutate(STUDYDAY = as.factor(STUDYDAY)) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE)) %>%
    dplyr::mutate(outlier = ifelse(is_outlier(SUM_SCORE), SUM_SCORE, as.numeric(NA))) %>%
    dplyr::arrange(STUDYDAY) %>%
  ggplot(aes(x = STUDYDAY, y = SUM_SCORE)) +
  geom_boxplot() +
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="Patient Symptom Score Sum",
       title=paste("Patient Symptom Score Sum Distribution by Virus:\n",
       label," Data",sep="")) +
    facet_wrap(~VIRUS)
}
```
  
```{r loading_data}
train_symptom_df <- read.table(TRAIN_SYMPTOM,header=T,sep="\t")
train_clinicl_df <- read.table(TRAIN_CLINICL,header=T,sep="\t")
phas1_clinicl_df <- read.table(PHAS1_CLINICL,header=T,sep="\t")
phas2_clinicl_df <- read.table(PHAS2_CLINICL,header=T,sep="\t")
phas3_clinicl_df <- read.table(PHAS3_CLINICL,header=T,sep="\t")
```

### Scan for Missing Data...

```{r scanning_for_missing_data}
sum(is.na(train_symptom_df)) #0
sum(is.na(train_clinicl_df)) #4318
sum(is.na(phas1_clinicl_df)) #60
sum(is.na(phas2_clinicl_df)) #88
sum(is.na(phas3_clinicl_df)) #62
```

### Fix Missing Data...

```{r fix_missing_training_clinical_data}
#### Training Clinical Data
train_clinicl_df %>% str()
train_clinicl_df %>% summary()
train_clinicl_df <- train_clinicl_df %>%
  dplyr::mutate(EARLYTX = ifelse(is.na(EARLYTX),0,1))
train_clinicl_df <- train_clinicl_df %>% 
  dplyr::mutate(SHAM = ifelse(is.na(SHAM),0,1))
```

```{r fix_missing_phase1_clinical_data}
#### Phase 1 Clinical Data
phas1_clinicl_df %>% str()
phas1_clinicl_df <- phas1_clinicl_df %>% 
  dplyr::mutate(EARLYTX = ifelse(is.na(EARLYTX),0,1))
phas1_clinicl_df <- phas1_clinicl_df %>% 
  dplyr::mutate(SHAM = ifelse(is.na(SHAM),0,1))
```

```{r fix_missing_phase2_clinical_data}
#### Phase 2 Clinical Data
phas2_clinicl_df %>% str()
phas2_clinicl_df <- phas2_clinicl_df %>%
  dplyr::mutate(EARLYTX = ifelse(is.na(EARLYTX),0,1))
phas2_clinicl_df <- phas2_clinicl_df %>%
  dplyr::mutate(SHAM = ifelse(is.na(SHAM),0,1))
```

```{r fix_missing_phase3_clinical_data}
#### Phase 3 Clinical Data
phas3_clinicl_df %>% str()
phas3_clinicl_df <- phas3_clinicl_df %>%
  dplyr::mutate(EARLYTX = ifelse(is.na(EARLYTX),0,1))
phas3_clinicl_df <- phas3_clinicl_df %>% 
  dplyr::mutate(SHAM = ifelse(is.na(SHAM),0,1))
```

### Find max/median log symptom score by time hours with study & virus summarizations

```{r lss_timehours_summary, echo=TRUE, results='asis'}
train_clinicl_df %>%
  dplyr::group_by(STUDYID) %>%
  dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
  dplyr::summarise(jackson_mean=mean(JACKSON_SCORE),
                  jackson_med=median(JACKSON_SCORE),
                  log_mean=mean(LOGSYMPTSCORE_SC3),
                  log_med=median(LOGSYMPTSCORE_SC3)) %>%
  pander(split.cells = 50, split.table = Inf)

train_clinicl_df %>%
  dplyr::group_by(STUDYID,TIMEHOURS) %>%
  dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
  dplyr::summarise(jackson_mean=mean(JACKSON_SCORE),
                  jackson_med=median(JACKSON_SCORE),
                  log_mean=mean(LOGSYMPTSCORE_SC3),
                  log_med=median(LOGSYMPTSCORE_SC3)) %>%
  pander(split.cells = 50, split.table = Inf)

train_clinicl_df %>%
  add_virus_col() %>%
  dplyr::group_by(VIRUS) %>%
  dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
  dplyr::summarise(jackson_mean=mean(JACKSON_SCORE),
                  jackson_med=median(JACKSON_SCORE),
                  log_mean=mean(LOGSYMPTSCORE_SC3),
                  log_med=median(LOGSYMPTSCORE_SC3)) %>%
  pander(split.cells = 50, split.table = Inf)

train_clinicl_df %>%
  add_virus_col() %>%
  dplyr::group_by(VIRUS,TIMEHOURS) %>%
  dplyr::mutate(JACKSON_SCORE = 2^LOGSYMPTSCORE_SC3) %>%
  dplyr::summarise(jackson_mean=mean(JACKSON_SCORE),
                  jackson_med=median(JACKSON_SCORE),
                  log_mean=mean(LOGSYMPTSCORE_SC3),
                  log_med=median(LOGSYMPTSCORE_SC3)) %>%
  pander(split.cells = 50, split.table = Inf)

z <- train_symptom_df %>% 
  dplyr::group_by(STUDYID) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE))
z %>% dplyr::ungroup() %>% dplyr::group_by(STUDYID) %>%
  dplyr::summarise(score_mean=mean(SUM_SCORE),
                   score_med=median(SUM_SCORE)) %>%
  pander(split.cells = 50, split.table = Inf)

z <- train_symptom_df %>% 
  dplyr::mutate(STUDYDAY = as.factor(STUDYDAY)) %>%
  dplyr::group_by(STUDYID,STUDYDAY) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE))
z %>% dplyr::ungroup() %>% dplyr::group_by(STUDYID,STUDYDAY) %>%
  dplyr::summarise(score_mean=mean(SUM_SCORE),
                   score_med=median(SUM_SCORE)) %>%
  pander(split.cells = 50, split.table = Inf)

z <- train_symptom_df %>% 
  add_virus_col() %>%
  dplyr::group_by(VIRUS) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE))
z %>% dplyr::ungroup() %>% dplyr::group_by(VIRUS) %>%
  dplyr::summarise(score_mean=mean(SUM_SCORE),
                   score_med=median(SUM_SCORE)) %>%
  pander(split.cells = 50, split.table = Inf)

z<- train_symptom_df %>% 
  add_virus_col() %>%
  dplyr::mutate(STUDYDAY = as.factor(STUDYDAY)) %>%
  dplyr::group_by(VIRUS,STUDYDAY) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE))
z %>% dplyr::ungroup() %>% dplyr::group_by(VIRUS,STUDYDAY) %>%
  dplyr::summarise(score_mean=mean(SUM_SCORE),
                   score_med=median(SUM_SCORE)) %>%
  pander(split.cells = 50, split.table = Inf)
```

```{r lss_time-day_distributions}
# train

train_clinicl_df %>%
  log_symptom_score_by_study("Train")
train_clinicl_df %>%
  add_virus_col() %>%
  log_symptom_score_by_virus("Train")

train_clinicl_df %>%
  log_symptom_score_by_timehours_by_study("Train")
train_clinicl_df %>%
  add_virus_col() %>%
  log_symptom_score_by_timehours_by_virus("Train")

train_symptom_df %>%
  symptom_score_by_studyday_by_study("Train")
train_symptom_df %>%
  add_virus_col() %>%
  symptom_score_by_studyday_by_virus("Train")
```

### Ensure all patients are between 15-60 yo

```{r age_summaries, echo=TRUE, results='asis'}
train_clinicl_df$AGE %>% summary()
phas1_clinicl_df$AGE %>% summary()
phas2_clinicl_df$AGE %>% summary()
phas3_clinicl_df$AGE %>% summary()
```

```{r age_distributions}
# train

train_clinicl_df %>%
  age_by_study("Train")
train_clinicl_df %>%
  add_virus_col() %>%
  age_by_virus("Train")

# phase 1 (0h)

phas1_clinicl_df %>%
  age_by_study("Phase 1 (0h)")
phas1_clinicl_df %>%
  add_virus_col() %>%
  age_by_virus("Phase 1 (0h)")

# phase 2

phas2_clinicl_df %>%
  age_by_study("Phase 2")
phas2_clinicl_df %>%
  add_virus_col() %>%
  age_by_virus("Phase 2")

# phase 3

phas3_clinicl_df %>%
  age_by_study("Phase 3")
phas3_clinicl_df %>%
  add_virus_col() %>%
  age_by_virus("Phase 3")

# 24 hour

phas2_clinicl_df %>%
  rbind(phas3_clinicl_df) %>%
  age_by_study("24h")
phas2_clinicl_df %>%
  rbind(phas3_clinicl_df) %>%
  add_virus_col() %>%
  age_by_virus("24h")
```

### Look at removing patients who report pre-challenge symptoms

```{r patients_with_pre-challenge_symptoms, echo=TRUE, results='asis'}
train_clinicl_df %>%
  dplyr::distinct(STUDYID,SUBJECTID) %>% dim()
train_clinicl_df %>%
  dplyr::filter(TIMEHOURS <= 0, LOGSYMPTSCORE_SC3 > 0) %>%
  dplyr::distinct(STUDYID,SUBJECTID) %>% dim()

train_symptom_df %>%
  dplyr::distinct(STUDYID,SUBJECTID) %>% dim()
train_symptom_df %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE)) %>%
  dplyr::filter(STUDYDAY <= 0,SUM_SCORE > 0) %>%
  dplyr::distinct(STUDYID,SUBJECTID) %>% dim()
```

### Could we use pre-challenge unreported symptoms as predictors?

```{r assess_pre-challenge_symptoms}
z <- train_symptom_df %>% add_virus_col()
z <- reshape2::melt(z[,c('STUDYDAY', 'STUDYID', 'VIRUS',
                               'SX_RUNNYNOSE',
                               'SX_COUGH',
                               'SX_HEADACHE',
                               'SX_MALAISE',
                               'SX_MYALGIA',
                               'SX_SNEEZE',
                               'SX_SORETHROAT',
                               'SX_STUFFYNOSE')],
                               id.vars = c('STUDYDAY','STUDYID','VIRUS'))
z %>%  
dplyr::filter(STUDYDAY <= 0,value > 0) %>%
dplyr::arrange(variable) %>%
  ggplot(aes(x = variable,fill=variable)) + 
  geom_bar(stat="count") +
  labs(x="Symptom",
       y="Pre-Challenge Symptom Counts",
       title="Pre-Challenge Symptom Counts\n
       by Study ID") +
  geom_text(stat="count",aes(label=..count..),vjust=-.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~STUDYID)

z %>%  
dplyr::filter(STUDYDAY <= 0,value > 0) %>%
dplyr::arrange(variable) %>%
  ggplot(aes(x = variable,fill=variable)) + 
  geom_bar(stat="count") +
  labs(x="Symptom",
       y="Pre-Challenge Symptom Counts",
       title="Pre-Challenge Symptom Counts\n
       by Virus") +
  geom_text(stat="count",aes(label=..count..),vjust=-.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~VIRUS)

z %>%  
dplyr::filter(STUDYDAY > 0,value > 0) %>%
dplyr::arrange(variable) %>%
  ggplot(aes(x = variable,fill=variable)) + 
  geom_bar(stat="count") +
  labs(x="Symptom",
       y="Post-Challenge Symptom Counts",
       title="Post-Challenge Symptom Counts\n
       by Study ID") +
  geom_text(stat="count",aes(label=..count..),vjust=-.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~STUDYID)

z %>%  
dplyr::filter(STUDYDAY > 0,value > 0) %>%
dplyr::arrange(variable) %>%
  ggplot(aes(x = variable,fill=variable)) + 
  geom_bar(stat="count") +
  labs(x="Symptom",
       y="Post-Challenge Symptom Counts",
       title="Post-Challenge Symptom Counts\n
       by Virus") +
  geom_text(stat="count",aes(label=..count..),vjust=-.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~VIRUS)
```

### What are the distributions of pre-challenge unreported symptoms?

```{r pre-challenge_unreported_distributions}

#Rhinovirus: Sneeze
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="Rhinovirus") %>%
  dplyr::select(STUDYDAY,SX_SNEEZE) %>%
  dplyr::arrange(SX_SNEEZE) %>%
  ggplot(aes(x=STUDYDAY,y=SX_SNEEZE,fill=SX_SNEEZE)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_SNEEZE (any severity)",
       title="Rhinovirus Study Day by SX_SNEEZE Count and Severity")
  
#H1N1: Cough
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="H1N1") %>%
  dplyr::select(STUDYDAY,SX_COUGH) %>%
  dplyr::arrange(SX_COUGH) %>%
  ggplot(aes(x=STUDYDAY,y=SX_COUGH,fill=SX_COUGH)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_COUGH (any severity)",
       title="H1N1 Study Day by SX_COUGH Count and Severity")

#RSV: Cough
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="RSV") %>%
  dplyr::select(STUDYDAY,SX_COUGH) %>%
  dplyr::arrange(SX_COUGH) %>%
  ggplot(aes(x=STUDYDAY,y=SX_COUGH,fill=SX_COUGH)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_COUGH (any severity)",
       title="RSV Study Day by SX_COUGH Count and Severity")

#RSV: Malaise
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="RSV") %>%
  dplyr::select(STUDYDAY,SX_MALAISE) %>%
  dplyr::arrange(SX_MALAISE) %>%
  ggplot(aes(x=STUDYDAY,y=SX_MALAISE,fill=SX_MALAISE)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_MALAISE (any severity)",
       title="RSV Study Day by SX_MALAISE Count and Severity")

#RSV: Myalgia
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="RSV") %>%
  dplyr::select(STUDYDAY,SX_MYALGIA) %>%
  dplyr::arrange(SX_MYALGIA) %>%
  ggplot(aes(x=STUDYDAY,y=SX_MYALGIA,fill=SX_MYALGIA)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_MYALGIA (any severity)",
       title="RSV Study Day by SX_MYALGIA Count and Severity")

#RSV: Sore Throat
train_symptom_df %>% add_virus_col() %>%
  dplyr::filter(VIRUS=="RSV") %>%
  dplyr::select(STUDYDAY,SX_SORETHROAT) %>%
  dplyr::arrange(SX_SORETHROAT) %>%
  ggplot(aes(x=STUDYDAY,y=SX_SORETHROAT,fill=SX_SORETHROAT)) +
  geom_bar(stat="identity") +
  labs(fill="Severity",
       x="Study Day",
       y="Number of Patients Reporting SX_SORETHROAT (any severity)",
       title="RSV Study Day by SX_SORETHROAT Count and Severity")

```

### What phenotypes (symptoms/symptom combinations) are not seen pre-challenge?

> Rhinovirus: Sneeze  
> H1N1: Cough  
> RSV: Cough  
> RSV: Malaise  
> RSV: Myalgia  
> RSV: Sore Throat  

### Look at removing patients who are asymptomatic for the entire study. Could we use them as shams?

```{r remove_asymptomatic_patients, echo=TRUE, results='asis'}
z <- train_symptom_df %>%
  dplyr::rowwise() %>%
  dplyr::mutate(SUM_SCORE = sum(SX_HEADACHE,
                                  SX_RUNNYNOSE,
                                  SX_COUGH,
                                  SX_MALAISE,
                                  SX_MYALGIA,
                                  SX_SNEEZE,
                                  SX_SORETHROAT,
                                  SX_STUFFYNOSE))
  z <- z %>% dplyr::ungroup() %>%
    dplyr::group_by(SUBJECTID,STUDYID) %>%
    dplyr::select(SUBJECTID,STUDYID,SUM_SCORE) %>%
  dplyr::summarize_each(dplyr::funs(sum)) %>%
  dplyr::filter(SUM_SCORE == 0)
  z %>% dim()
  z %>%
  pander(split.cells = 50, split.table = Inf)
  
  z <- train_clinicl_df %>%
    dplyr::group_by(SUBJECTID,STUDYID) %>%
    dplyr::select(SUBJECTID,STUDYID,LOGSYMPTSCORE_SC3) %>%
  dplyr::summarize_each(dplyr::funs(sum)) %>%
  dplyr::filter(LOGSYMPTSCORE_SC3 == 0)
  z %>% dim()
  z %>%
  pander(split.cells = 50, split.table = Inf)
```

### Explore ID's of CEL files

```{r CEL_file_names}
train_clinicl_df %>% 
  dplyr::mutate(CEL_num = as.numeric(CEL)) %>%
  dplyr::arrange(CEL_num) %>%
  ggplot(aes(x=STUDYID,y=CEL_num)) +
  geom_point() +   
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="Study ID",
       y="CEL Filename (converted to numeric for ranking)",
       title="CEL Filename by Study ID")
```

### Explore early TX
